---
title: "Data 607_Week 3 HW_ER"
author: "Euclides Rodriguez"
date: "2/20/2021"
output: html_document
---
```{r include=FALSE}
library(tidyverse)
```
  
#### Week 3 Excercise 1 
Filtering a data frame:

```{r}
df <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv')

df %>% 
  filter(str_detect(Major, "DATA|STATISTICS"))
```

  
#### Week 3 Excercise 2 
Combining a string:
```{r}
str_c(c("bell pepper","bilberry","blackberry","blood orange","blueberry","cantaloupe","chili pepper","cloudberry", "olive","salal berry"), collapse=", ")

```

  
#### Week 3 Excercise 3
Describe the following expressions in words:<p>
* (.)\1\1<br />
<p style="color:blue"> One character followed by 2 matching characters<p>  
* "(.)(.)\\\2\\\1"<br />
<p style="color:blue"> Double back slash indicates for use with str_view. Two single character group followed by matching of 2nd group letter followed by last character matching the 1st<p>  
* (..)\1<br />
<p style="color:blue"> Two character group followed by identical two character.<p>  
* "(.).\\\1.\\\1"<br />
<p style="color:blue"> Double back slash indicates for use with str_view. One single character group, followed any character followed by 1st group letter follow by any letter followed by first group letter.<p>  
* "(.)(.)(.).*\\\3\\\2\\\1"<br />
<p style="color:blue"> Double back slash indicates for use with str_view. Three single character groups followed by any character followed by matching of third group letter followed bye 2nd group letter followed by 1st group letter.<p>  

  
#### Week 3 Excercise 4 
Construct regular expressions to match words that:<p>
* Start and end with the same character.<br />
<p style="color:blue"> (.).*\1<p>

* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)<br />
<p style="color:blue">(..).*\1<p>
* Contain one letter repeated in at least three places eleven" contains three "e"s.)<br />
<p style="color:blue">(.).\*\1.\*\1<p>


